\documentclass[paper=a4, fontsize=11pt]{scrartcl}
\usepackage[T1]{fontenc}
%\usepackage{fourier}

\usepackage[english]{babel}
\usepackage[protrusion=true,expansion=true]{microtype}	
\usepackage{amsmath,amsfonts,amsthm} % Math packages
\usepackage[pdftex]{graphicx}	
\usepackage{url}
\usepackage{makecell,pict2e}
\usepackage{comment}
\usepackage{color}

\renewcommand\theadfont{\large}
\newcommand{\gt}[1]{{\color{blue}#1}}
\newcommand{\red}[1]{{\color{red}#1}}

\renewcommand\theadfont{\large}

%%% Custom sectioning
\usepackage{sectsty}
\allsectionsfont{\normalfont\scshape}

\includecomment{answers}
% \excludecomment{answers}

\usepackage[margin=2cm]{geometry}
\setlength{\topmargin}{-2.cm}
\setlength{\headheight}{1cm}

%%% Custom headers/footers (fancyhdr package)
\usepackage{fancyhdr}
\pagestyle{fancyplain}
\fancyhead{}
\fancyfoot[L]{}
\fancyfoot[C]{}
\fancyfoot[R]{\thepage}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\setlength{\headheight}{13.6pt}

%%% Equation and float numbering
\numberwithin{equation}{section}
\numberwithin{figure}{section}
\numberwithin{table}{section}


%%% Maketitle metadata
\newcommand{\horrule}[1]{\rule{\linewidth}{#1}}
\newcommand{\vek}[1]{\mbox{\boldmath $  #1$}}
\newcommand{\ex}[1]{\ensuremath {\mathbb{E}} \left[ #1 \right]}
\newcommand{\var}[1]{\ensuremath{{\rm var}\left[ #1 \right]}}

\title{\usefont{OT1}{bch}{b}{n} \normalfont \normalsize \textsc{SOR3012:
Stochastic Processes} \\ [25pt] \horrule{0.5pt} \\[0.4cm] 
\huge Systems of queues \\
\horrule{2pt} \\[0.25cm]
}
\author{ \normalfont
\normalsize
        Gareth Tribello \\[-3pt] \normalsize
        \today
}
\date{}

\begin{document}
\maketitle

For this project I would like you to consider the way patients are treated in a hospital.

\begin{enumerate}
\item Patients arrive into accident and emergency with a rate $\lambda$ per minute.
\item Each patient waits to be seen by a triage nurse who then takes takes an average of $\frac{1}{2\lambda}$ minutes to deal with each patient.
\item One third of the patients that the triage nurse sees are then sent to clinic A where they wait for an appointment with a doctor.  On average a doctor in clinic A takes $\frac{3}{2\lambda}$ minutes to deal with a patient.
\item The remaining two thirds of the customers, i.e. those that were not sent to clinic A, are sent to clinic B where doctors take an average of $\frac{4}{3\lambda}$ minutes to deal with them. 
\end{enumerate}

Use the information in the description above to write a program that simulates the passage of patients through the hospital.  Use your program to estimate the average 
ammount of time a customer spends in hospital from their arrival in accident and emergency to their departure from clinic A/B after having been seen by a doctor.  Calculate
appropriate confidence limits on your estimate of this mean.

\end{document}
