\documentclass[a4paper]{article} 
\usepackage{tcolorbox}
\usepackage{amsmath,amsfonts,amsthm}
\tcbuselibrary{skins}
\usepackage{graphicx}
\usepackage{hyperref}

\makeatletter
\def\cornell{\@ifnextchar[{\@with}{\@without}}
\def\@with[#1]#2#3{
\begin{tcolorbox}[enhanced,colback=white!15,colframe=white,colupper=gray]
\begin{tcolorbox}[enhanced,colback=gray,colframe=black,fonttitle=\large\bfseries\sffamily,sidebyside, nobeforeafter,colupper=black,
righthand width=14cm,
opacityframe=0,opacityback=0.3,opacitybacktitle=1, opacitytext=1,
segmentation style={black!55,solid,opacity=0,line width=1pt},
title=#1
]
%\begin{tcolorbox}[colback=red!05,colframe=red!25,sidebyside align=top,
%width=\textwidth,nobeforeafter]#2\end{tcolorbox}%
%\tcblower
%\sffamily
%\begin{tcolorbox}[colback=blue!05,width=\textwidth]
% #3
%\end{tcolorbox}
\Huge {\bf  #2}
\tcblower
#3
\end{tcolorbox}
\end{tcolorbox}
}
\makeatother

\title{
\vspace{-3em}
\begin{tcolorbox}
\Huge\sffamily AMA4004 Statistical mechanics: Thermodynamics potentials   
\end{tcolorbox}
\vspace{-3em}
}

\date{}

%\usepackage{background}
%\SetBgScale{1}
%\SetBgAngle{0}
%\SetBgColor{red}
%\SetBgContents{\rule[0em]{4pt}{\textheight}}
%\SetBgHshift{-2.3cm}
%\SetBgVshift{0cm}
\usepackage{lipsum}% just to generate filler text for the example
\usepackage[margin=1.5cm]{geometry}
\usepackage{lipsum}% just to generate dummy text for the example


\begin{document}
\maketitle

{\bf This assignment contains no hard parts and as such if this report is handed in for the portfolio the maximum mark you can get for it is 8/12.} 

In order to do this exercise you will need to watch the following videos:

\begin{itemize}
\item Thermodynamic potentials: \href{http://gtribello.github.io/mathNET/reservoirs-and-thermodynamic-potentials-video.html}{http://gtribello.github.io/mathNET/reservoirs-and-thermodynamic-potentials-video.html}
\item Maxwell Relations: \href{http://gtribello.github.io/mathNET/maxwell-relations-video.html}{http://gtribello.github.io/mathNET/maxwell-relations-video.html}
\end{itemize}

It will probably be useful for you to also look at:

\href{http://gtribello.github.io/mathNET/maxwell-relations-exercise.html}{http://gtribello.github.io/mathNET/maxwell-relations-exercise.html}

You will be asked to prepare a short (1-page max) report on one of the following thermodynamic potentials:

\begin{itemize}
\item The internal energy
\item The Enthalpy
\item The Helmholtz free energy
\item The Gibbs free energy
\item The grand potential
\item A thermodynamic potential that is suitable for studying nucleation
\end{itemize}

Please typeset your report using \LaTeX. Define all mathematical symbols that you used and provide written justification for the steps you take in the derivations.  For example when you use results that only hold for exact differentials you should explain that the quantity under study is an exact differential.  I do not want a report that consists of equations only with no explanations.  In your report you will be expected to:

\begin{itemize}
\item Explain how the thermodynamic potential is defined and make reference to the nature of the reservoir that is in contact with the system.
\item Derive an expression for the differential i.e. some expression similar to $\textrm{d}E = T\textrm{d}S - P\textrm{d}V$
\item Show which thermodynamic variables can be calculated from partial derivatives of the thermodynamic potential and how.
\item Derive a Gibbs relation from the definition of the thermodynamic potential. 
\end{itemize}

You should consult multiple sources (not just the material on the mathNET website).  Please ensure you properly reference all the derivations you have included.

\end{document}